---
title: "Football Prediction"
author: "Yağmur Topçu"
date: "2026-04-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```


## Introduction

This project aims to predict football match outcomes (Home Win, Draw, Away Win) 
using statistical features such as team form, betting odds, and expected goals (xG).

The goal is to compare different machine learning models and evaluate their performance.

## Data

The dataset consists of multiple seasons of premier league football matches, including:
- Match results
- Betting odds
- Team performance metrics

Additional advanced metrics such as xG and PPDA are incorporated.

```{r}
library(here)
library(caret)
```


```{r}
source(here("scripts", "01_data_loading.R"))
source(here("scripts", "02_feature_engineering.R"))
source(here("scripts", "03_modeling.R"))
source(here("scripts", "04_evaluation.R"))
```
## Feature Engineering

Several features were created to capture team performance:
- Rolling form (last 5 matches)
- Goal difference trends
- Betting implied probabilities
- xG-based attacking and defensive strength

## Modeling

The following models were implemented:
- Multinomial Logistic Regression
- Random Forest
Models were trained on historical seasons and tested on the latest season.

```{r}
ggplot(results, aes(x = reorder(Model, Accuracy), y = Accuracy)) +
  geom_col(fill = "steelblue") +
  coord_flip() +
  labs(
    title = "Model Performance Comparison",
    x = "Model",
    y = "Accuracy"
  ) +
  theme_minimal()
```

```{r}
print(results)
```

```{r}
caret::confusionMatrix(as.factor(pred_rf), as.factor(test$FTR))

```


## Insights

- Betting odds are the strongest predictors of match outcomes
- xG features improve model performance
- Random Forest achieves the highest accuracy
- Draw outcomes remain the most difficult to predict










